Neutron Star Mergers and their Impact on Second Generation Star Formation in the Early Universe

### 2023-08-31 NSM_Skinner_Wise/referee/referee_response.tex

\begin{tcolorbox}[colback={lightgray}]
    8)      Purple lines in Figures 3, 5, and 8 show “Hyp.”. What is the meaning of Hyp.?
\end{tcolorbox}

We apologize for that confusion. "Hyp." stood for "hypernova" which is how the original Pop III star ends its life. To keep things simple, we have changed that in the plots to say "OG" to make it clear that those lines correspond to the original run with no NSM.

\begin{tcolorbox}[colback={lightgray}]
    9)      The authors show the coexistence of Pop II and III stars in the first galaxies. Recent theoretical studies showed similar results (Riaz et al. 2022, ApJL, 937, L6; Yajima et al., arXiv2211.12970). It would be good to discuss the mass fraction and distribution of Pop III stars in the first galaxies with the previous works. 
\end{tcolorbox}

\textbf{DS Notes:}
\begin{figure}[h]
    \centering
    \subfloat[\centering Energy Variation]{{\includegraphics[width=0.4\textwidth]{plots/eng_pop3_ratio.pdf} }}%
    \qquad
    \subfloat[\centering Delay Time Variation]{{\includegraphics[width=0.4\textwidth]{plots/time_pop3_ratio.pdf} }}%
    \caption{The mass fraction of Pop III stars to total stellar mass versus the total stellar mass colored by redshift.}%
    \label{fig:ratio}%
\end{figure}

\begin{figure}[h]
    \centering
    \subfloat[\centering Energy Variation]{{\includegraphics[width=0.4\textwidth]{plots/eng_stellar_mass.pdf} }}%
    \qquad
    \subfloat[\centering Delay Time Variation]{{\includegraphics[width=0.4\textwidth]{plots/time_stellar_mass.pdf} }}%
    \caption{The total stellar mass of both Pop III and Pop II stars as a function of redshift. The solid lines are the total stellar masses of Pop III stars, and the dashed lines are total stellar masses of Pop II stars.}%
    \label{fig:ratio}%
\end{figure}

\begin{itemize}
    \item Riaz+22: compare with their Figure 1 and 3
    \begin{itemize}
        \item Re. Figure 1: Looking at the redshift range between 10 and 20, and comparing with their halo masses between $10^7 - 10^9 M_\odot$ since this is the closest out halo gets to by the end of the simulation, we do see much higher stellar masses compared to them. For example, at z = 10, they see $10^3 - 10^5 M_\odot$ for Pop II stars and $10^1 - 10^2 M_\odot$ for Pop III stars. For us, at z = 10 we see a few $\times 10^4 M_\odot$ for Pop II stars (which aligns with their results), and $~10^{3} M_{\odot}$ for Pop III stars, which is certainly larger than Riaz+22. However, they do give some generous ranges for the Pop II and Pop III stellar masses, so our results don't seem to be completely different from what they are saying.
        \item Re. Figure 3: This figure is very similar to the one from Yajima+23. We do see similar ratios overall (see Figure \ref{fig:ratio} above).
    \end{itemize}
    \item Yajima+23: compare with their Figure 4, third bullet point in their conclusion
    \begin{itemize}
        \item Re: Figure 4. I have made this same plot for our work. We have slightly lower values compared to Yajima. For all variation runs, for a stellar mass of $~10^{4.5} M_\odot$, we see a Pop III mass ratio of about $~10^{-1.5}$, as compared to Yajima's results of $~10^{-1}$. They have considerable scatter for their z = 10, 11, and 12 data. Certainly we see the same trend as them, as stellar mass increases, the ratio of Pop III stellar mass to total stellar mass decreases. We do see similar values at our maximum stellar mass value of $~10^{5.5} M_\odot$ having a mass ratio of $~10^{-2.25}$. An important thing to consider in this comparison is the much higher halo mass of their main halos. They look at a halo that grows to $5 \times 10^{11} M_\odot$ by z = 9.5, whereas our halo gets to $~2 \times 10^{8} M_\odot$ at z = 10 (see Figure \ref{fig:ratio} above).
        \item Based on these results, we certainly agree with their third bullet point. In general, the mass fraction of Pop III stars is less than 0.01 for halos with stellar masses $> 10^{4} M_\odot$ 
    \end{itemize}
\end{itemize}

### 2023-08-31 NSM_Skinner_Wise/referee/referee_response.tex

\begin{tcolorbox}[colback={lightgray}]
    9)      The authors show the coexistence of Pop II and III stars in the first galaxies. Recent theoretical studies showed similar results (Riaz et al. 2022, ApJL, 937, L6; Yajima et al., arXiv2211.12970). It would be good to discuss the mass fraction and distribution of Pop III stars in the first galaxies with the previous works. 
\end{tcolorbox}

\textbf{DS Notes:}
\begin{figure}[h]
    \centering
    \subfloat[\centering Energy Variation]{{\includegraphics[width=0.4\textwidth]{plots/eng_pop3_ratio.pdf} }}%
    \qquad
    \subfloat[\centering Delay Time Variation]{{\includegraphics[width=0.4\textwidth]{plots/time_pop3_ratio.pdf} }}%
    \caption{The mass fraction of Pop III stars to total stellar mass versus the total stellar mass colored by redshift.}%
    \label{fig:ratio}%
\end{figure}

\begin{figure}[h]
    \centering
    \subfloat[\centering Energy Variation]{{\includegraphics[width=0.4\textwidth]{plots/eng_stellar_mass.pdf} }}%
    \qquad
    \subfloat[\centering Delay Time Variation]{{\includegraphics[width=0.4\textwidth]{plots/time_stellar_mass.pdf} }}%
    \caption{The total stellar mass of both Pop III and Pop II stars as a function of redshift. The solid lines are the total stellar masses of Pop III stars, and the dashed lines are total stellar masses of Pop II stars.}%
    \label{fig:ratio}%
\end{figure}

\begin{itemize}
    \item Riaz+22: compare with their Figure 1 and 3 (https://iopscience.iop.org/article/10.3847/2041-8213/ac8ea6/meta)
    \begin{itemize}
        \item Re. Figure 1: Looking at the redshift range between 10 and 20, and comparing with their halo masses between $10^7 - 10^9 M_\odot$ since this is the closest out halo gets to by the end of the simulation, we do see much higher stellar masses compared to them. For example, at z = 10, they see $10^3 - 10^5 M_\odot$ for Pop II stars and $10^1 - 10^2 M_\odot$ for Pop III stars. For us, at z = 10 we see a few $\times 10^4 M_\odot$ for Pop II stars (which aligns with their results), and $~10^{3} M_{\odot}$ for Pop III stars, which is certainly larger than Riaz+22. However, they do give some generous ranges for the Pop II and Pop III stellar masses, so our results don't seem to be completely different from what they are saying.
        \item Re. Figure 3: This figure is very similar to the one from Yajima+23. We do see similar ratios overall (see Figure \ref{fig:ratio} above).
    \end{itemize}
    \item Yajima+23: compare with their Figure 4, third bullet point in their conclusion (https://arxiv.org/pdf/2211.12970.pdf)
    \begin{itemize}
        \item Re: Figure 4. I have made this same plot for our work. We have slightly lower values compared to Yajima. For all variation runs, for a stellar mass of $~10^{4.5} M_\odot$, we see a Pop III mass ratio of about $~10^{-1.5}$, as compared to Yajima's results of $~10^{-1}$. They have considerable scatter for their z = 10, 11, and 12 data. Certainly we see the same trend as them, as stellar mass increases, the ratio of Pop III stellar mass to total stellar mass decreases. We do see similar values at our maximum stellar mass value of $~10^{5.5} M_\odot$ having a mass ratio of $~10^{-2.25}$. An important thing to consider in this comparison is the much higher halo mass of their main halos. They look at a halo that grows to $5 \times 10^{11} M_\odot$ by z = 9.5, whereas our halo gets to $~2 \times 10^{8} M_\odot$ at z = 10 (see Figure \ref{fig:ratio} above).
        \item Based on these results, we certainly agree with their third bullet point. In general, the mass fraction of Pop III stars is less than 0.01 for halos with stellar masses $> 10^{4} M_\odot$ 
    \end{itemize}
\end{itemize}

### 2023-09-13 NSM_Skinner_Wise/NSM_mixing.tex
Adjusting text to reflect referee report response.

% Abstract of the paper
\begin{abstract}
	The exact evolution of elements in the universe, from primordial hydrogen and helium to heavier elements like gold and platinum produced via the r-process, is still under scrutiny. The supernova deaths of the very first stars to form in the universe led to the enrichment of their local environments with new metals, and can leave behind neutron stars as remnants. These remnants can end up in binary systems with other neutron stars, and eventually merge, allowing for the r-process to occur. In this work, we study the scenario where a single neutron star merger (NSM) enriches a halo early in its evolution to understand the impact on the second generation of stars and their metal abundances. We perform a suite of high resolution cosmological zoom-in simulations using \enzo{} where we have implemented a new NSM model varying the explosion energy and the delay time. In general, a NSM leads to a significant r-process enhancement in the second generation of stars. A high explosion energy leads to almost all enhanced r-process stars being highly enhanced, while a lower explosion energy leads to a higher mass fraction of stars being r-process enhanced, but not as many being highly enhanced. When a NSM has a short delay time, there is a higher mass fraction of stars being r-process enhanced, but a smaller mass fraction being highly enhanced compared to longer delay times. This work represents a stepping stone towards understanding how NSMs impact their environments and metal abundances of descendant generations of stars.
\end{abstract}

@@

The motivation for this work comes from the r-process enriched stars in Ret II, an ultrafaint dwarf (UFD) galaxy which serves as an ideal laboratory for studying early universe star formation. These systems typically form most of their stars early in the universe, before reionization was able to suppress star formation \citep{Weisz14}. This means that these stars form in the first couple hundred million years after the Big Bang and haven't been affected by recent mergers or enriched by recent star formation. Ret II was detected by the Dark Energy Survey \citep{Bechtol15}, whose stars have since been extensively studied \citep{Simon15, Roederer16, Ji16, Ji22}. Recently, \citet{Ji22} have found that $\sim 72\%$ of the stars in Ret II are r-process enriched, whose source could have been a prompt NSM. Since this galaxy is so isolated, it serves as an ideal case study for simulating this scenario: namely, what happens if a single NSM, produced from the remnants of Pop III stars, occurs within a galaxy in the early universe?

@@

\subsection{Neutron star merger model and simulation suite} \label{sec:NSM_model}
We implement a new NSM model within \enzo{} to control the delay time and the explosion energy of the event, modelled as a single star particle. Upon completion of the original simulation, a single Pop III star is chosen as the candidate NSM event. The chosen Pop III star must be between $20 \leq M / M_{\odot} \leq 30$ so that when this system is split into a symmetric binary system, the initial masses of the component Pop III stars is between $10 \leq M / M_{\odot} \leq 15$, corresponding to the SN models from \citet{Heger10} which will produce a NS as a remnant. We use their model C as a point of reference since the explosion energy, $0.9 \times 10^{51}$ erg, is the closest to the CCSN explosion energy in \enzo{}. The simulation is restarted at the output where the Pop III star forms, indicating which Pop III star and what NSM parameters should be included. We chose the first allowed Pop III star that forms in the simulation to become the source of the NSM event in order to allow for the maximum amount of time between the NSM event and for the second generation of stars to form. The Pop III star that we chose has an initial mass of $28.7 M_{\odot}$ that was randomly selected from an IMF described in \citet{Skinner20}. The chosen Pop III star is then converted into a symmetric Pop III binary system with component masses of $\sim 14.35 M_{\odot}$, but is still modelled as a single star particle with an ionizing hydrogen photon rate of $2.9 \times 10^{48} \mathrm{s}^{-1}$, which is equal to the sum of the two individual stellar photon rates from \citet{Schaerer02}.

@@

A second explosion goes off with $E_{\textrm{NSM}}$ after a time $t_{\textrm{merge}}$, with the parameters chosen in a particular run given the choices in Table \ref{tab:NSM_params}. The final remnant particle is then converted to an inert particle with a mass equal to the sum of the remnant masses. We do not consider any EM radiation from the NSs nor any kilonova effects of a NSM for simplicity; we only follow the mass ejected from this event. We track the r-process material in a passive scalar field which does not affect metal cooling rates. For this proof of concept, the total ejecta mass is set to $0.01 \Ms$, but in practice, due to the Cartesian discretization of a sphere in of Enzo, the actual mass ejected is $0.0169 M_{\odot}$.

@@

A Pop II star forms using the same criteria, except without the molecular hydrogen fraction requirement, and with a metallicity greater than the above value. A Pop II particle represents a cluster of Pop II stars with a minimum cluster mass of 1000 \Ms. There are instances when a Pop II star will form below the minimum mass, in which case the star accretes mass without producing feedback until it reaches that minimum mass. Both Pop III and Pop II stars are assigned a metallicity based on the metallicity fraction of the cell that the particle forms in. Because Pop II stars form as star clusters, the exact number of Pop II star particles that form is not a good indicator of how productive an environment is at forming Pop II stars. Rather, the total Pop II stellar mass is the indicator that will be presented throughout this work.

@@

\begin{figure} 
	\includegraphics[width=\columnwidth]{plots/eng_stellar_dist.pdf}
	\caption[Radial distribution of Pop II stars for each explosion energy variation run.]{The fraction of Pop II stars that are within some radius at the $z = 10$ for runs D, Fiducial and E, and the original run. The dashed lines indicate the half-mass radius, which are 331.2 pc, 37.1 pc, 65.3 pc, and 50.6 pc for the original run, run D, run Fiducial, and run E, respectively. The virial radius at this time is 1.6 kpc, 1.73 kpc, 1.72 kpc, and 1.73 kpc for the original run, run D, run Fiducial, and run E, respectively.}
	\label{fig:eng_stellar_dist}
\end{figure}

@@

@@

This work represents some of the first simulations on the chemical and dynamical impacts of a NSM, with various parameters, in a cosmological setting, and thus there are some difficulties that come with this kind of study. The first is that the small sample size prevents us from saying anything conclusive about the NSM parameters in general. While we see some trends across the data, the inherent randomness across simulations leads to different amounts of star formation at different times and locations. On top of that, because we randomly sample the Pop III IMF, if two simulations are looking very similar and both go on to form a Pop III star in a similar location, one can be sampled at a low mass and the other at a very high mass, and affect the surrounding region in completely different ways. What may seem like a small difference in the Pop III stellar mass can go on to have impactful consequences on future star formation. This means that it can be difficult to pick apart why certain stars form where and at what times. We stress that this is merely a proof-of-concept study to show that we can study the impacts of NSMs in a cosmological setting to present meaningful results that can enhance our understanding of r-process enrichment in the early universe.

>>>>

% Abstract of the paper
\begin{abstract}
	The exact evolution of elements in the universe, from primordial hydrogen and helium to heavier elements like gold and platinum produced via the r-process, is still under scrutiny. The supernova deaths of the very first stars to form in the universe led to the enrichment of their local environments with new metals, and can leave behind neutron stars as remnants. These remnants can end up in binary systems with other neutron stars, and eventually merge, allowing for the r-process to occur. In this work, we study the scenario where a single neutron star merger (NSM) enriches a halo early in its evolution to understand the impact on the second generation of stars and their metal abundances. We perform a suite of high resolution cosmological zoom-in simulations using \enzo{} where we have implemented a new NSM model varying the explosion energy and the delay time. In general, a NSM leads to a significant r-process enhancement in the second generation of stars. \textbf{A high explosion energy leads to a Pop II mass fraction of 72\% being highly enhanced with r-process elements, while a lower explosion energy leads to a Pop II mass fraction of 80\% being r-process enhanced, but only 14\% being highly enhanced. When the NSM has a short delay time of 10 Myr, only 5\% of the mass fraction of Pop II stars is highly r-process enhanced, while 64\% of the mass fraction of Pop II stars is highly r-process enhanced for the longest delay time of 100 Myr.} This work represents a stepping stone towards understanding how NSMs impact their environments and metal abundances of descendant generations of stars.
\end{abstract}

@@

The motivation for this work comes from the r-process enriched stars in Ret II, an ultrafaint dwarf (UFD) galaxy which serves as an ideal laboratory for studying early universe star formation. These systems typically form most of their stars early in the universe, before reionization was able to suppress star formation \citep{Weisz14}. This means that these stars form in the first couple hundred million years after the Big Bang and haven't been affected by recent mergers or enriched by recent star formation. Ret II was detected by the Dark Energy Survey \citep{Bechtol15}, whose stars have since been extensively studied \citep{Simon15, Roederer16, Ji16, Ji22}. Recently, \citet{Ji22} have found that $\sim 72\%$ of the stars in Ret II are r-process enriched, whose source could have been a prompt NSM. Since this galaxy is so isolated, it serves as an ideal case study for simulating this scenario: namely, what happens if a single NSM, produced from the remnants of Pop III stars, occurs within a galaxy in the early universe? \textbf{This type of scenario is particularly unique due to the chemical impact a NSM could have on the star formation history of a galaxy. If a NSM did occur early in the history of Ret II, the r-process enriched stars are a special and direct connection to previous generations of stars.}

@@

\subsection{Neutron star merger model and simulation suite} \label{sec:NSM_model}
\textbf{To explore how a NSM descendent from Pop III stars affects the r-process enrichment of Pop II stars, w}e implement a new NSM model within \enzo{} to control the delay time and the explosion energy of the event, modelled as a single star particle. Upon completion of the original simulation, a single Pop III star is chosen as the candidate NSM event. The chosen Pop III star must be between $20 \leq M / M_{\odot} \leq 30$ so that when this system is split into a symmetric binary system, the initial masses of the component Pop III stars is between $10 \leq M / M_{\odot} \leq 15$, corresponding to the SN models from \citet{Heger10} which will produce a NS as a remnant. We use their model C as a point of reference since the explosion energy, $0.9 \times 10^{51}$ erg, is the closest to the CCSN explosion energy in \enzo{}. The simulation is restarted at the output where the Pop III star forms, indicating which Pop III star and what NSM parameters should be included. We chose the first allowed Pop III star that forms in the simulation to become the source of the NSM event in order to allow for the maximum amount of time between the NSM event and for the second generation of stars to form. The Pop III star that we chose has an initial mass of $28.7 M_{\odot}$ that was randomly selected from an IMF described in \citet{Skinner20}. The chosen Pop III star is then converted into a symmetric Pop III binary system with component masses of $\sim 14.35 M_{\odot}$, but is still modelled as a single star particle with an ionizing hydrogen photon rate of $2.9 \times 10^{48} \mathrm{s}^{-1}$, which is equal to the sum of the two individual stellar photon rates from \citet{Schaerer02}.

@@

A second explosion goes off with $E_{\textrm{NSM}}$ after a time $t_{\textrm{merge}}$, with the parameters chosen in a particular run given the choices in Table \ref{tab:NSM_params}. The final remnant particle is then converted to an inert particle with a mass equal to the sum of the remnant masses. We do not consider any EM radiation from the NSs nor any kilonova effects of a NSM for simplicity; we only follow the mass ejected from this event. We track the r-process material in a passive scalar field which does not affect metal cooling rates. For this proof of concept, the total ejecta mass is set to $0.01 \Ms$, but in practice, due to the Cartesian discretization of a sphere in of Enzo, the actual mass ejected is $0.0169 M_{\odot}$. \textbf{The delay times chosen for these runs are on the shorter end of timescales for the merging of two NSs to take place, but for this proof of concept, we wanted to explore this particular situation where a NSM occurs early in the star formation history of the galaxy. For a galaxy like Ret II, a NSM would have had to occur within 500 Myr to allow for ample time for thorough mixing to take place within the halo \citep{Simon23}. Delay times of 10, 100, and 300 Myr are short, but are not completely out of the realm of possibilities, and allow us to see how a NSM affects subsequent star formation at this early point in time.}

@@

A Pop II star forms using the same criteria, except without the molecular hydrogen fraction requirement, and with a metallicity greater than the above value. A Pop II particle represents a cluster of Pop II stars with a minimum cluster mass of 1000 \Ms. There are instances when a Pop II star will form below the minimum mass, in which case the star accretes mass without producing feedback until it reaches that minimum mass. Both Pop III and Pop II stars are assigned a metallicity based on the metallicity fraction of the cell that the particle forms in. Because Pop II stars form as star clusters, the exact number of Pop II star particles that form is not a good indicator of how productive an environment is at forming Pop II stars. Rather, the total Pop II stellar mass is the indicator that will be presented throughout this work. \textbf{We do not take into account the effects of Pop III or Pop II binary systems due to the limiting resolution of our simulations.}

@@

\begin{figure} 
	\includegraphics[width=\columnwidth]{plots/eng_stellar_dist.pdf}
	\caption[Radial distribution of Pop II stars for each explosion energy variation run.]{The fraction of Pop II stars that are within some radius at the $z = 10$ for runs D, Fiducial and E, and the original \textbf{(OG)} run. The dashed lines indicate the half-mass radius, which are 331.2 pc, 37.1 pc, 65.3 pc, and 50.6 pc for the original run, run D, run Fiducial, and run E, respectively. The virial radius at this time is 1.6 kpc, 1.73 kpc, 1.72 kpc, and 1.73 kpc for the original run, run D, run Fiducial, and run E, respectively.}
	\label{fig:eng_stellar_dist}
\end{figure}

@@

\textbf{}

@@

\textbf{For all runs, we see similar Pop III stellar masses as compared to some recent studies. \citet{Riaz22} studied the Pop III contribution to high redshift galaxies for a range of halo masses. They find that a halo with a mass of $10^7$ \Ms has a Pop III mass fraction of $\sim 10^{-1}$ at $z \approx 12$. For our runs, at a similar time and halo mass, we also see a Pop III mass fraction of $\sim 10^{-1}$. By $z = 10$, we see $\sim 10^{4}$ \Ms of Pop II stars and $\sim 10^{3}$ \Ms for Pop III stars, as compared to $10^3 - 10^5$ \Ms of Pop II stars and  $10^1 - 10^2$ \Ms of Pop III stars found by \citet{Riaz22}. While we do see a slightly higher value of Pop III stellar masses as compared to them, our overall mass fractions are similar. \citet{Yajima23} studied the first galaxies and their Pop III contributions in order to compare with results from JWST. They also look at the Pop III mass fraction (see their Figure 5) as compared to the total stellar mass. They report a much broader range of values for different redshifts in general. For a stellar mass of $10^{4.5}$ \Ms, they find a Pop III mass fraction of $\sim 10^{-0.75}$, while we see a mass fraction for that stellar mass closer to $10^{-1.5}$ \Ms. However, we do see a similar trend, that as stellar mass increases, the Pop III mass fraction begins to decrease. At the end of our simulations when our halo has a stellar mass of $\sim 10^{5.5}$ \Ms, we see a Pop III mass fraction of $\sim 10^{-2.25}$, which is similar to \citet{Yajima23}. Based on these results, we agree with their conclusion that for halos with stellar masses of $\sim 10^{7}$ \Ms, the Pop III mass fraction is small, at $< 0.01$ }

This work represents some of the first simulations on the chemical and dynamical impacts of a NSM, with various parameters, in a cosmological setting, and thus there are some difficulties that come with this kind of study. The first is that the small sample size prevents us from saying anything conclusive about the NSM parameters in general. While we see some trends across the data, the inherent randomness across simulations leads to different amounts of star formation at different times and locations. On top of that, because we randomly sample the Pop III IMF, if two simulations are looking very similar and both go on to form a Pop III star in a similar location, one can be sampled at a low mass and the other at a very high mass, and affect the surrounding region in completely different ways. What may seem like a small difference in the Pop III stellar mass can go on to have impactful consequences on future star formation. This means that it can be difficult to pick apart why certain stars form where and at what times. We stress that this is merely a proof-of-concept study to show that we can study the impacts of NSMs in a cosmological setting to present meaningful results that can enhance our understanding of r-process enrichment in the early universe. \textbf{Studies are currently underway to try to separate the source of differences between the simulations, and are outside the scope of this paper.}
